---
title: "Fialko-Shvets-lab4"
output: html_document
date: "2022-12-21"
---

## Team:

-   Shvets Anastasia

-   Fialko Yaryna

### Generating Data

$$
{a_k} := {kln(k^2n + π)},  k≥1
$$

Where ${x} := x - [x]$ is a fractional part of a number x, n - id number

$$
   \mathscr{N}(\mu_1,\sigma_1^2) \qquad
$$

$$
{x_k}= Ф^{-1}(a_k), k= 1,...,100
$$

$$
   \mathscr{N}(\mu_2,\sigma_2^2) \qquad
$$

$$
{y_l} = Ф^{-1}(a_{l+100}), l = 1,...,50
$$

Where $Ф$ - is the cumulative distribution function of N(0, 1) and $Ф_{-1}$ is inverse
```{r}
n =  10
calculate_a <- function(k) {
   return(k*log((k^2)*n + pi))
}

a_list <- list()
for (i in 1:100) {
  a_list <- append(a_list, calculate_a(i))
}
ay_list <- list()
for (i in 1:50) {
  ay_list <- append(ay_list, calculate_a(i+100))
}

print(ay_list)

```

Problem 1:

$H_0\,: \mu_1 = \mu_2 \quad \textrm{vs} \quad H_1\,: \mu ≠ \mu_0$

$\sigma_1^2 = \sigma_2^2 = 1$

```{r}

```

**We have applied a Z-Test**

A z-test is a statistical test used to determine whether two population means are different when the variances are known and the sample size is large.

### Problem 2:

$H_0\,: \sigma_1^2 = \sigma_2^2 \quad \textrm{vs} \quad H_1\,: \sigma_1^2 > \sigma_2^2$

$\mu_1$ and $\mu_2$ are unknown

### Problem 3:

Using Kolmogorov--Smirnov test in R, check if
(a) ${(xk)}_{k=1}^{100}$ are normally distributed (with parameters calculated from the sample);

(b) ${(|xk|)}_{k=1}^{100}$ are exponentially distributed with λ = 1;
(c) ${(xk)}_{k=1}^{100}$ and \${y_l}\_{l=1}\^{50}\$ have the same distributions.




